# Results {#sec:results}

## Real-Game Illustration of STRAIN

<!-- ```{r} -->
<!-- #| label: fig_field -->
<!-- #| fig.cap: "A display of the player tracking data for a play during the Las Vegas Raiders (defense, in black) versus Denver Broncos (offense, in orange) NFL game on October 17, 2021. Raiders DE Maxx Crosby is highlighted in blue. For each Raiders defender, the point size indicates their individual STRAIN value, with larger points suggesting larger STRAIN. Snapshots are captured at 1, 2, 3, and 4 seconds after the ball snap." -->
<!-- #| fig.width: 6.6 -->
<!-- #| fig.height: 4.4 -->
<!-- source(here("scripts", "01_fig_field.R")) -->
<!-- fig_field -->
<!-- ``` -->

<!-- Figure \ref{fig:fig_field} shows an updated version of Figure \ref{fig:fig_field_raw}, with the point size for each Las Vegas defender corresponding to the estimated STRAIN in the selected frames as the play progresses. -->

To illustrate our proposed metric STRAIN for pass rush evaluation, we use the example play from the Las Vegas Raiders and Denver Broncos game as mentioned in Section \ref{sec:data}.
Figure \ref{fig:fig_features_combined} (left) is a line graph showing how Crosby's distance from the quarterback, velocity, and STRAIN change continuously throughout the play.
In addition, Figure C.1 of the Supplement presents the locations of every Las Vegas and Denver player
on the field from this play at 1, 2, 3 and 4 seconds after the ball snap, with the point size for each Las Vegas defender corresponding to the estimated STRAIN at the selected frames.

We observe that for the first two seconds, Crosby is being blocked by a Denver offensive lineman and unable to get close to the quarterback, hence the corresponding STRAIN values are virtually zero. 
Suddenly at around three seconds after the snap, the STRAIN for Crosby starts to increase after the Raiders defensive end is freed up and charges towards the QB. 
At 4 seconds after the snap, Crosby's STRAIN is 2.30, which means at his current moving rate, it will take Crosby about 0.43 (1/2.30) seconds to make the distance between him and the quarterback 0 (i.e. essentially sack the quarterback).
This matches well with the final outcome of the play, as the sack takes place at the very last frame (4.4 seconds) where the estimated STRAIN for Crosby reaches its peak at 3.96.

Moreover, we clearly see the interactions between the features for Maxx Crosby from Figure \ref{fig:fig_features_combined} (left).
Here, a higher STRAIN generally corresponds to faster moving rate towards the quarterback.
STRAIN also increases as the distance between the pass rusher and quarterback is being reduced.
Both of these relationships suggest an overall successful pass rush by Maxx Crosby.
It is also notable that Crosby, who is credited with a sack, generates more STRAIN than other pass rushers during this play.

In contrast, Figure \ref{fig:fig_features_combined} (right) shows the feature curves for an unsuccessful pass rush attempt by the Raiders defense from the same game.
In this [play](https://www.nfl.com/videos/teddy-bridgewater-courtland-sutton-barely-miss-out-on-chance-at-62-yard-td-bomb), Broncos QB Teddy Bridgewater is well-protected by the offensive line and is able to release a long pass to a receiver.
We see that the STRAIN generated by Crosby is relatively small over the course of this play, compared to the previous play which results in a sack.

```{r}
#| label: fig_features_combined
#| fig.cap: "Changes in STRAIN, velocity, and distance from the quarterback for Maxx Crosby over the course of a successful pass rush play that results in a sack (left) and an unsuccessful pass rush play which results in a long pass attempt (right). The STRAIN for other pass rushers throughout each play is also displayed."
#| fig.width: 6.1
#| fig.height: 3.4
source(here("scripts", "15_fig_features_combined.R"))
fig_features_combined_bw
```

<!-- [^unsuccess]: Game footage of this play is available at https://www.nfl.com/videos/teddy-bridgewater-courtland-sutton-barely-miss-out-on-chance-at-62-yard-td-bomb -->


<!-- ```{r} -->
<!-- #| label: fig_crosby_curves_unsucess -->
<!-- #| fig.cap: "Changes in STRAIN, velocity, and distance from the quarterback for Maxx Crosby over the course of an unsuccessful pass rush play. The STRAIN for other pass rushers throughout this play is also displayed." -->
<!-- #| fig.width: 4.8 -->
<!-- #| fig.height: 2.7 -->
<!-- source(here("scripts", "09_fig_play_unsuccess.R")) -->
<!-- fig_crosby_curves_unsucess -->
<!-- ``` -->


<!-- ```{r} -->
<!-- #| label: fig_features_combined -->
<!-- #| fig.cap: "Changes in STRAIN, velocity, and distance from the quarterback for Maxx Crosby over the course of a successful pass rush play that results in a sack (left) and an unsuccessful pass rush play which results in a long pass attempt (right). The STRAIN for other pass rushers throughout each play is also displayed." -->
<!-- #| fig.width: 6 -->
<!-- #| fig.height: 3.6 -->
<!-- source(here("scripts", "15_fig_features_combined.R")) -->
<!-- fig_features_combined_bw -->
<!-- ``` -->

<!-- Figure \ref{fig:fig_features_combined} -->



## Positional STRAIN Curves {#sec:positional}

```{r}
#| label: fig_pos_curves
#| fig.cap: "STRAIN curves for different positions, along with the average STRAIN curve for all players without accounting for position. Edge rushers (outside linebackers and defensive ends) tend to generate more STRAIN than interior rushers (defensive tackles and nose tackles)."
#| fig.width: 5
#| fig.height: 3
source(here("scripts", "03_fig_pos_curves.R"))
fig_pos_curves_bw
```

```{r}
#| label: fig_outcome_curves
#| fig.cap: "Positional STRAIN curves by play outcome (sack, hit, hurry, and none)."
#| fig.width: 6.5
#| fig.height: 2.7
source(here("scripts", "08_fig_outcome_curves.R"))
fig_outcome_curves_bw
```

<!-- 
too wordy and confusing imo...
For each frame, this curve is calculated by averaging across all players of a particular position (where position is defined by the NFL) and all plays that lasted at least as long as the particular frame of interest meaning that the sample size used to compute these curves diminishes for frames that are chronologically later than frames at the beginning of the play. -->

Figure \ref{fig:fig_pos_curves} displays the average STRAIN by position for the first 40 frames (4 seconds) after the snap.
For each position, the curve is based on the average of the observed STRAIN at each frame across all plays and players within the position.
We observe a clear difference in STRAIN between edge rushers (outside linebackers and defensive ends) and interior linemen (defensive tackles and nose tackles).
Specifically, edge rushers have higher STRAIN than interior linemen on average, as they are more easily able to approach the quarterback on the edge of the pocket versus, for instance, a nose tackle attacking the line head on.

On average, STRAIN appears to increase for the first 0.5 seconds of a play, followed by a decline in the next second.
STRAIN then increases again until around 2.5--3 seconds after the snap before trending down towards the end of the play.
In context, this reflects the actions that a pass rusher initially moves towards the quarterback, but is then stopped by the offensive line while the quarterback drops back.
When the quarterback stops dropping, the rusher closes the gap and increases STRAIN, before slowing down later on.

Further, Figure \ref{fig:fig_outcome_curves} shows the positional STRAIN curves by whether a play's outcome is a hit, sack, hurry, or none of those. 
We clearly see that players tend to generate more STRAIN when a play ends in a hit, sack, or hurry, compared to no outcome.
Within the three pressure metrics, it is not surprising that a sack corresponds to the highest amount of STRAIN, followed by a hit and hurry.

<!-- ## Ranking the Best Pass Rushers {#sec:strainbar} -->

<!-- Since STRAIN is observed continuously for every play in our data, this allows us to aggregate across all frames played and compute the average STRAIN for NFL pass rushers over the course of the eight-week sample size, as discussed in Section \ref{sec:defn}. -->
<!-- Based on the clearly distinct patterns for different positions as previously observed, we evaluate interior pass rushers (nose tackles and defensive tackles) separately from edge rushers (outside linebackers and defensive ends). -->
<!-- Tables \ref{tab:edge} and \ref{tab:interior} are leaderboards for the NFL's best edge and interior rushers (with at least 100 plays) rated by the average STRAIN across all frames for the first eight weeks of the 2021 regular season. -->
<!-- The tables also consist of the total number of hits, hurries, and sacks (determined from PFF scouting data) for each defender. -->

<!-- <!-- Our results show many true positives with no surprises. -->
<!-- <!-- Rashan Gary, a young rusher with huge potential, ranks first among all pass rushers according to our metric over the period of time observed. -->

<!-- Our results are mostly consistent with conventional rankings of rushers. -->
<!-- Notably, Myles Garrett and TJ Watt are widely recognized as top-tier edge rushers and both show up in our top edge rusher list; whereas Aaron Donald, who is undoubtedly the best interior defender in football, appears at the top of our interior rusher rankings. -->
<!-- Moreover, our leaderboards largely match the rankings of experts in the field. -->
<!-- For instance, there is considerable overlap between our lists and PFF's edge [@Monson2022nfl] and interior [@Linsey2022nfl] rusher rankings released after the 2021 season. -->
<!-- This ultimately lends credibility to our proposed metric STRAIN as a measure of pass rushing effectiveness. -->

<!-- \begin{table} -->
<!-- \caption{Top 15 edge rushers (with at least 100 snaps played) according to the average STRAIN across all frames played. \label{tab:edge}} -->
<!-- \centering -->
<!-- \begin{tabular}{rlllrrrrr} -->
<!-- \hline -->
<!-- Rank & Player & Team & Position & Snaps & Hits & Hurries & Sacks & $\overline{\text{STRAIN}}$ \\ -->
<!-- \hline -->
<!-- 1 & Rashan Gary & GB & OLB & 176 & 10 & 25 & 5 & 2.82 \\  -->
<!-- 2 & Leonard Floyd & LA & OLB & 185 & 2 & 25 & 8 & 2.80 \\  -->
<!-- 3 & Justin Houston & BAL & OLB & 132 & 8 & 8 & 4 & 2.78 \\  -->
<!-- 4 & Myles Garrett & CLE & DE & 197 & 9 & 29 & 12 & 2.75 \\  -->
<!-- 5 & Von Miller & DEN & OLB & 145 & 4 & 21 & 5 & 2.75 \\  -->
<!-- 6 & T.J. Watt & PIT & OLB & 147 & 6 & 9 & 8 & 2.71 \\  -->
<!-- 7 & Yannick Ngakoue & LV & DE & 175 & 6 & 20 & 4 & 2.70 \\  -->
<!-- 8 & Alex Highsmith & PIT & OLB & 129 & 4 & 7 & 2 & 2.65 \\  -->
<!-- 9 & Preston Smith & GB & OLB & 124 & 4 & 8 & 2 & 2.61 \\  -->
<!-- 10 & Randy Gregory & DAL & DE & 134 & 7 & 19 & 5 & 2.58 \\  -->
<!-- 11 & Joey Bosa & LAC & OLB & 160 & 5 & 21 & 4 & 2.58 \\  -->
<!-- 12 & Darrell Taylor & SEA & DE & 107 & 5 & 9 & 3 & 2.57 \\  -->
<!-- 13 & Josh Sweat & PHI & DE & 159 & 4 & 14 & 5 & 2.57 \\  -->
<!-- 14 & Maxx Crosby & LV & DE & 198 & 12 & 30 & 7 & 2.56 \\  -->
<!-- 15 & Markus Golden & ARI & OLB & 164 & 5 & 14 & 5 & 2.50 \\  -->
<!-- \hline -->
<!-- \end{tabular} -->
<!-- \end{table} -->

<!-- \begin{table} -->
<!-- \caption{Top 15 interior rushers (with at least 100 snaps played) according to the average STRAIN across all frames played. \label{tab:interior}} -->
<!-- \centering -->
<!-- \begin{tabular}{rlllrrrrr} -->
<!-- \hline -->
<!-- Rank & Player & Team & Position & Snaps & Hits & Hurries & Sacks & $\overline{\text{STRAIN}}$ \\ -->
<!-- \hline -->
<!-- 1 & Aaron Donald & LA & DT & 239 & 8 & 24 & 6 & 1.67 \\  -->
<!-- 2 & Solomon Thomas & LV & DT & 115 & 7 & 11 & 3 & 1.51 \\  -->
<!-- 3 & Quinton Jefferson & LV & DT & 144 & 6 & 8 & 3 & 1.46 \\  -->
<!-- 4 & Chris Jones & KC & DT & 139 & 3 & 18 & 3 & 1.42 \\  -->
<!-- 5 & DeForest Buckner & IND & DT & 198 & 4 & 18 & 4 & 1.26 \\  -->
<!-- 6 & Cameron Heyward & PIT & DT & 188 & 2 & 22 & 3 & 1.25 \\  -->
<!-- 7 & Javon Hargrave & PHI & DT & 156 & 6 & 15 & 5 & 1.24 \\  -->
<!-- 8 & Jerry Tillery & LAC & DT & 171 & 4 & 7 & 3 & 1.16 \\  -->
<!-- 9 & Ed Oliver & BUF & DT & 133 & 4 & 12 & 1 & 1.15 \\  -->
<!-- 10 & Osa Odighizuwa & DAL & DT & 162 & 3 & 18 & 3 & 1.13 \\  -->
<!-- 11 & Greg Gaines & LA & NT & 111 & 2 & 13 & 2 & 1.11 \\  -->
<!-- 12 & Leonard Williams & NYG & DT & 226 & 4 & 14 & 6 & 1.03 \\  -->
<!-- 13 & Christian Barmore & NE & DT & 166 & 5 & 17 & 1 & 1.02 \\  -->
<!-- 14 & Vita Vea & TB & NT & 184 & 6 & 12 & 1 & 1.01 \\  -->
<!-- 15 & B.J. Hill & CIN & DT & 123 & 3 & 4 & 3 & 0.96 \\  -->
<!-- \hline -->
<!-- \end{tabular} -->
<!-- \end{table} -->


## Statistical Properties of STRAIN {#sec:statprop}

Next, we examine different statistical properties of our proposed metric STRAIN.
Our focus here is to understand the stability and predictability of STRAIN, and we pose the following question:
How much does our metric vary from week to week?
In other words, is previous performance predictive of future performance based on our metric?
Below, we attempt the answer these questions, and the following results are for pass rushers with at least 100 snaps played during the first eight weeks of the 2021 NFL regular season.

We first investigate the predictability of STRAIN as a measure of pressure.
In particular, we look at how well our metric correlates with a simple measure of pressure rate, defined as the total hits, sacks and hurries per snap.
Figure \ref{fig:fig_cor_pressure} is a scatterplot of average STRAIN and pressure rate over the course the provided eight-game sample, which reveals a fairly strong correlation ($r =0.6255$) between the quantities.
Hence, defenders with high STRAIN values also tend to generate more pressure toward the quarterback.
Furthermore, the average STRAIN for the first four weeks of the 2021 season is more predictive of the last four weeks' pressure rate ($r=0.3217$) than the first four weeks' pressure rate ($r=0.0965$), as illustrated in Figure \ref{fig:fig_predictability}.

We also analyze the stability of STRAIN over time by comparing the average STRAIN across all frames played for the first and last four weeks of the 2021 NFL regular season.
It is apparent from Figure \ref{fig:fig_stability} that there is a strong positive correlation $(r = 0.8545)$ for this relationship.
This means that STRAIN is a highly stable football metric over the provided eight-week time window, and pass rushers appear to carry their STRAIN values with them from week to week.
Overall, STRAIN performs well in both explaining defensive pressure on the field and predicting future performance of pass rushers.


```{r}
#| label: fig_cor_pressure
#| fig.cap: "Relationship between average STRAIN and pressure rate (total hits, sacks, and hurries per snap) over the first eight weeks of the 2021 NFL season. There is a moderately strong association between average STRAIN and pressure rate ($r =0.6255$). Results shown here are for pass rushers with at least 100 snaps played over the eight-week data."
#| fig.width: 4
#| fig.height: 3
source(here("scripts", "04_fig_cor_pressure.R"))
fig_cor_pressure_bw
```

```{r}
#| label: fig_predictability
#| fig.cap: "Relationships between pressure rate for the last 4 weeks of the 2021 NFL season and first four weeks' pressure rate (left) and average STRAIN (right). STRAIN is more predictive ($r=0.3217$) of future pressure rate than previous pressure rate itself ($r=0.0965$). Results shown here are for pass rushers with at least 100 snaps played over the eight-week data."
#| fig.width: 6.5
#| fig.height: 3
#| fig.show: "hold"
source(here("scripts", "07_fig_predictability.R"))
fig_predictability_bw
```

```{r}
#| label: fig_stability
#| fig.cap: "Relationship between average STRAIN for the last and first four weeks of the 2021 NFL season. A strong linear correlation ($r = 0.8545$) demonstrates that STRAIN is a highly stable metric over time. Results shown here are for pass rushers with at least 100 snaps played over the eight-week data."
#| fig.width: 4
#| fig.height: 3
source(here("scripts", "05_fig_stability.R"))
fig_stability_bw
```

## Multilevel Model Results

The results of fitting the multilevel model described in Section \ref{sec:multilevel} is displayed in Table \ref{tab:fixedeff}.
First, we investigate the fixed effects terms of this model.
It appears that that the average STRAIN decreases by 0.7366 (95% CI [-0.7875, -0.6858]) for every additional blocker in the offensive unit, after accounting for other covariates.
In other words, NFL pass rushers tend to generate more pressure when facing fewer number of blockers, which makes intuitive sense.

As for play context, defensive pressure appears to increase when more yardage is required for the offense to reach a first down and for later plays within a set of down, when all other predictors are held constant.
In particular, every extra yard in the distance needed for a first down is associated with a 0.0491 increase (95% CI [0.0426, 0.0555]) in the average STRAIN.
Relative to first down situations, higher amount of pressure seem to happen in plays that come after (second, third, and fourth downs).
In addition, we have insufficient evidence for a relationship between average STRAIN and the current yardline on the field for the offensive team.

Moreover, we observe statistically significant differences among the pass rush and pass block positions in most cases, while controlling for other variables.
The coefficient estimates for the pass rush position terms reveal that players with positions closer to the line of scrimmage (defensive tackles and nose tackles) tend to generate less STRAIN (relative to defensive ends) than those lining up further back, which is consistent with our results in Section \ref{sec:positional}.
For blockers, centers are those that absorb the most pressure on average, more than guards and other offensive positions (compared to the baseline level, tackles).

Next, we examine the intraclass correlation coefficients (ICC) for the four different groupings
in our multilevel model.
This descriptive statistic illustrates the proportion variance explained between the group terms in comparison to the residual variance.
While the residual variance is unsurprisingly the highest value (0.9260), between the player and team factors, we observe the largest ICC for pass rushers (0.0365), followed by the offensive (0.0143) and defensive teams (0.0134), and finally the blocker (0.0098).
This emphasizes how STRAIN is mostly attributable to pass rushers, but is necessary to adjust for opposition and other factors.

<!-- reference levels: DE (rush), T (block) -->

<!-- \begin{table} -->
<!-- \caption{Fixed-effects coefficient estimates for the multilevel model for average STRAIN. Note that the reference pass rusher level is defensive end (DE), denoted R:DE; and the reference pass blocker level is tackle (T), denoted B:T. \label{tab:fixedeff}} -->
<!-- \centering -->
<!-- \begin{tabular}{lrrrr} -->
<!-- \hline -->
<!-- & estimate & se & $t$-statistic & $p$-value \\  -->
<!-- \hline -->
<!-- Intercept & 0.378 & 0.009 & 43.310 & 0.000 \\  -->
<!-- Length of play & -0.042 & 0.001 & -38.214 & 0.000 \\ -->
<!-- Number of blockers & -0.064 & 0.003 & -24.788 & 0.000 \\ -->
<!-- $I_{\text{\{R:DT\}}}$ & -0.081 & 0.008 & -10.340 & 0.000 \\ -->
<!-- $I_{\text{\{R:interior\}}}$ & 0.058 & 0.011 & 5.332 & 0.000 \\ -->
<!-- $I_{\text{\{R:NT\}}}$ & -0.106 & 0.012 & -9.174 & 0.000 \\ -->
<!-- $I_{\text{\{R:OLB\}}}$ & 0.062 & 0.008 & 7.695 & 0.000 \\ -->
<!-- $I_{\text{\{R:secondary\}}}$ & 0.145 & 0.012 & 12.328 & 0.000 \\ -->
<!-- $I_{\text{\{B:C\}}}$ & -0.023 & 0.006 & -3.625 & 0.000 \\ -->
<!-- $I_{\text{\{B:G\}}}$ & -0.011 & 0.005 & -2.290 & 0.023 \\  -->
<!-- $I_{\text{\{B:other\}}}$ & -0.013 & 0.007 & -1.911 & 0.056 \\ -->
<!-- \hline -->
<!-- \end{tabular} -->
<!-- \end{table} -->

<!-- (Intercept) & 0.378 & 0.009 & 43.310 & 0.000 \\  -->
<!--   play\_len & -0.004 & 0.000 & -38.214 & 0.000 \\  -->
<!--   rush\_posDT & -0.081 & 0.008 & -10.340 & 0.000 \\  -->
<!--   rush\_posinteriorLB & 0.058 & 0.011 & 5.332 & 0.000 \\  -->
<!--   rush\_posNT & -0.106 & 0.012 & -9.174 & 0.000 \\  -->
<!--   rush\_posOLB & 0.062 & 0.008 & 7.695 & 0.000 \\  -->
<!--   rush\_possecondary & 0.145 & 0.012 & 12.328 & 0.000 \\  -->
<!--   block\_posC & -0.023 & 0.006 & -3.625 & 0.000 \\  -->
<!--   block\_posG & -0.011 & 0.005 & -2.290 & 0.023 \\  -->
<!--   block\_posother & -0.013 & 0.007 & -1.911 & 0.056 \\  -->
<!--   n\_blockers & -0.064 & 0.003 & -24.788 & 0.000 \\  -->

\begin{table}
\caption{Fixed effects coefficient estimates for the multilevel model for average STRAIN. Note that the reference down level is first down, the reference pass rusher level is defensive end (DE), denoted R:DE; the reference pass blocker level is tackle (T), denoted B:T. \label{tab:fixedeff}}
\centering
\begin{tabular}{lrrrr}
\hline
& estimate & se & $t$-statistic & $p$-value \\ 
\hline
Intercept & 1.8004 & 0.0991 & 18.1723 & 0.0000 \\ 
Number of blockers & -0.7366 & 0.0259 & -28.4153 & 0.0000 \\
Yards to go & 0.0491 & 0.0033 & 14.9013 & 0.0000 \\
Current yardline & 0.0001 & 0.0005 & 0.1844 & 0.8537 \\
$I_{\text{\{2nd down\}}}$ & 0.5143 & 0.0290 & 17.7439 & 0.0000 \\
$I_{\text{\{3rd down\}}}$ & 0.9176 & 0.0324 & 28.3613 & 0.0000 \\
$I_{\text{\{4th down\}}}$ & 0.4649 & 0.0767 & 6.0637 & 0.0000 \\
$I_{\text{\{2pt conversion\}}}$ & -0.1384 & 0.1797 & -0.7699 & 0.4413 \\
$I_{\text{\{R:DT\}}}$ & -0.7166 & 0.0678 & -10.5658 & 0.0000 \\
$I_{\text{\{R:interior\}}}$ & 0.5611 & 0.0999 & 5.6143 & 0.0000 \\
$I_{\text{\{R:NT\}}}$ & -0.9426 & 0.1021 & -9.2344 & 0.0000 \\
$I_{\text{\{R:OLB\}}}$ & 0.5574 & 0.0726 & 7.6742 & 0.0000 \\
$I_{\text{\{R:secondary\}}}$ & 1.4252 & 0.1114 & 12.7899 & 0.0000 \\
$I_{\text{\{B:C\}}}$ & -0.2983 & 0.0594 & -5.0237 & 0.0000 \\
$I_{\text{\{B:G\}}}$ & -0.1546 & 0.0471 & -3.2847 & 0.0000 \\ 
$I_{\text{\{B:other\}}}$ & -0.1260 & 0.0646 & -1.9508 & 0.0514 \\
\hline
\end{tabular}
\end{table}


<!-- \begin{table}[ht] -->
<!-- \caption{Intraclass correlation coefficients for the multilevel model for average STRAIN. \label{tab:icc}} -->
<!-- \centering -->
<!-- \begin{tabular}{lr} -->
<!-- \hline -->
<!-- Group & ICC \\  -->
<!-- \hline -->
<!-- Pass rusher & 0.057 \\  -->
<!-- Pass blocker & 0.012 \\  -->
<!-- Quarterback & 0.023 \\  -->
<!-- Residual & 0.907 \\  -->
<!-- \hline -->
<!-- \end{tabular} -->
<!-- \end{table} -->


<!-- \begin{table}[ht] -->
<!-- \caption{Intraclass correlation coefficients for the multilevel model for average STRAIN. \label{tab:icc}} -->
<!-- \centering -->
<!-- \begin{tabular}{lr} -->
<!-- \hline -->
<!-- Group & ICC \\  -->
<!-- \hline -->
<!-- Pass rusher & 0.0365 \\  -->
<!-- Pass blocker & 0.0098 \\  -->
<!-- Defensive team & 0.0134 \\  -->
<!-- Offensive team & 0.0143 \\ -->
<!-- Residual & 0.9260 \\  -->
<!-- \hline -->
<!-- \end{tabular} -->
<!-- \end{table} -->

<!-- \begin{table} -->
<!-- \caption{Intraclass correlation coefficients for the multilevel model for average STRAIN. \label{tab:icc}} -->
<!-- \centering -->
<!-- \begin{tabular}{cccccc} -->
<!-- \hline -->
<!-- \text{ } & Pass rusher & Pass blocker & Defensive team & Offensive team & Residual \\  -->
<!-- \hline -->
<!-- ICC & 0.0365 & 0.0098 & 0.0134 & 0.0143 & 0.9260 \\  -->
<!-- \hline -->
<!-- \end{tabular} -->
<!-- \end{table} -->

```{r}
#| label: fig_rankings_boot
#| fig.cap: "Distributions of player effects (obtained from 1,000 bootstrap samples) for top 10 pass rushers in each position (defensive end, outside linebacker, defensive tackle, nose tackle) by median varying intercept."
#| fig.width: 6.3
#| fig.height: 6
source(here("scripts", "10_multilevel_mod.R"))
fig_rankings_boot
```

<!-- We display the random intercepts for pass rusher in Table [INSERT TABLE REFERENCE] alongside their unadjusted STRAIN values.  -->

<!-- Our results are mostly consistent with conventional rankings of rushers, with Myles Garrett (defensive end), T.J. Watt (outside linebacker), Aaron Donald (defensive tackle), and Vita Vea (nose tackle) as the first-ranked pass rusher for their respective position. -->
<!-- Moreover, our leaderboards largely match the rankings of experts in the field. -->
<!-- For instance, there is considerable overlap between our lists and PFF's edge [@Monson2022nfl] and interior [@Linsey2022nfl] rusher rankings released after the 2021 season. -->
<!-- This ultimately lends credibility to our proposed metric STRAIN as a measure of pass rushing effectiveness. -->

Aside from adjusting for opposition, our multilevel model allows us to provide some notion of uncertainty about the pass rusher's effect on the play level STRAIN.
Figure \ref{fig:fig_rankings_boot} displays the varying intercept distributions for the top ten pass rushers for each position.
Our rankings based on the random intercepts are mostly consistent with conventional rankings of rushers, with Myles Garrett (defensive end), T.J. Watt (outside linebacker), Aaron Donald (defensive tackle), and Vita Vea (nose tackle) as the first-ranked pass rusher for their respective position.
Garrett and Watt are widely recognized as top-tier edge rushers, whereas Donald is undoubtedly the best interior defender in football.
Moreover, our leaderboards largely match the rankings of experts in the field.
For instance, there is considerable overlap between our lists and PFF's edge [@Monson2022nfl] and interior [@Linsey2022nfl] rusher rankings released after the 2021 season.
This ultimately lends credibility to our proposed metric STRAIN as a measure of pass rushing effectiveness.
Additionally, we observe different levels of variability across these top defenders, with only a subset of players having intercept distributions strictly above zero.
This is not necessarily surprising given the limited sample of data.

<!-- [TODO: edit the below? McKinley still there but we did control for team effects] -->

<!-- Additionally, we are ignoring correlated performance between defensive linemen on the same team. -->
<!-- For instance, the top two defensive ends, Myles Garrett and Takkarist McKinley (see the top panel of Figure \ref{fig:fig_rankings_boot}), were teammates on the Cleveland Browns during this period of time. -->
<!-- McKinley is a surprising name on our list and we suspect that his high rank is mostly due to playing on the same defensive unit with Garrett, who is well-known as an elite pass rusher. -->
<!-- We leave the exploration of adjusting for teammates for future work. -->