# Results {#sec:results}

## Example play

```{r}
#| label: fig_field
#| fig.cap: "snapshots"
#| fig.width: 6.6
#| fig.height: 4.4
source(here("scripts", "01_fig_field.R"))
fig_field
```


Figure \ref{fig:fig_field} shows snapshots

```{r}
#| label: fig_crosby_curves
#| fig.cap: "crosby"
#| fig.width: 4.8
#| fig.height: 2.7
source(here("scripts", "02_fig_crosby_curves.R"))
fig_crosby_curves
```

Figure \ref{fig:fig_crosby_curves} shows Crosby throughout the play

## Positional STRAIN curves

```{r}
#| label: fig_pos_curves
#| fig.cap: "STRAIN curves for different positions"
#| fig.width: 4.5
#| fig.height: 3
source(here("scripts", "03_fig_pos_curves.R"))
fig_pos_curves
```

Figure \ref{fig:fig_pos_curves} shows pos curves

## Ranking the best pass-rushers

\begin{table}
\caption{Top 15 edge rushers (with at least 100 snaps played) according to the average STRAIN across all frames played. \label{tab:edge}}
\centering
\begin{tabular}{rrlllrr}
\hline
 Rank & Player & Team & Position & Snaps & Average STRAIN \\ 
\hline
1 & Rashan Gary & GB & OLB & 176 & 2.82 \\ 
2 & Leonard Floyd & LA & OLB & 185 & 2.80 \\ 
3 & Justin Houston & BAL & OLB & 132 & 2.78 \\ 
4 & Myles Garrett & CLE & DE & 197 & 2.75 \\ 
5 & Von Miller & DEN & OLB & 145 & 2.75 \\ 
6 & T.J. Watt & PIT & OLB & 147 & 2.71 \\ 
7 & Yannick Ngakoue & LV & DE & 175 & 2.70 \\ 
8 & Alex Highsmith & PIT & OLB & 129 & 2.65 \\ 
9 & Preston Smith & GB & OLB & 124 & 2.61 \\ 
10 & Randy Gregory & DAL & DE & 134 & 2.58 \\ 
11 & Joey Bosa & LAC & OLB & 160 & 2.58 \\ 
12 & Darrell Taylor & SEA & DE & 107 & 2.57 \\ 
13 & Josh Sweat & PHI & DE & 159 & 2.57 \\ 
14 & Maxx Crosby & LV & DE & 198 & 2.56 \\ 
15 & Markus Golden & ARI & OLB & 164 & 2.50 \\ 
\hline
\end{tabular}
\end{table}



\begin{table}
\caption{Top 15 interior rushers (with at least 100 snaps played) according to the average STRAIN across all frames played. \label{tab:interior}}
\centering
\begin{tabular}{rlllrr}
\hline
Rank & Player & Team & Position & Snaps & Average STRAIN \\ 
\hline
1 & Aaron Donald & LA & DT & 239 & 1.67 \\ 
2 & Solomon Thomas & LV & DT & 115 & 1.51 \\ 
3 & Quinton Jefferson & LV & DT & 144 & 1.46 \\ 
4 & Chris Jones & KC & DT & 139 & 1.42 \\ 
5 & DeForest Buckner & IND & DT & 198 & 1.26 \\ 
6 & Cameron Heyward & PIT & DT & 188 & 1.25 \\ 
7 & Javon Hargrave & PHI & DT & 156 & 1.24 \\ 
8 & Jerry Tillery & LAC & DT & 171 & 1.16 \\ 
9 & Ed Oliver & BUF & DT & 133 & 1.15 \\ 
10 & Osa Odighizuwa & DAL & DT & 162 & 1.13 \\ 
11 & Greg Gaines & LA & NT & 111 & 1.11 \\ 
12 & Leonard Williams & NYG & DT & 226 & 1.03 \\ 
13 & Christian Barmore & NE & DT & 166 & 1.02 \\ 
14 & Vita Vea & TB & NT & 184 & 1.01 \\ 
15 & B.J. Hill & CIN & DT & 123 & 0.96 \\ 
\hline
\end{tabular}
\end{table}


Table \ref{tab:edge}

Table \ref{tab:interior}






## Corr with pressure rate = (hits + sacks + hurries) per snap

```{r}
#| label: fig_cor_pressure
#| fig.cap: "Cor with"
#| fig.width: 4
#| fig.height: 3
source(here("scripts", "04_fig_cor_pressure.R"))
fig_cor_pressure
```

Figure \ref{fig:fig_cor_pressure} shows cor with pressure

Fairly strong correlation (r = 0.6254943) between average STRAIN & hurries + sacks + hits per snaps

## Statistical Properties of STRAIN

From @Franks2016Meta (github https://github.com/afranks86/meta-analytics)

<!--

Discrimination: Does the metric reliably differentiate between players?

Independence: Does the metric provide new information?

-->

Stability: Does the metric measure a quantity which is stable over time?  (section 5.1 of openWAR has a similar idea)

```{r}
#| label: fig_stability
#| fig.cap: "Metric stability"
#| fig.width: 4
#| fig.height: 3
source(here("scripts", "05_fig_stability.R"))
fig_stability
```

0.8544865

Figure \ref{fig:fig_stability} shows

Average STRAIN across all frames played in first and last 4 weeks

Strong correlation shows great metric stability
